---
title: "Anomaly_inspection"
author: "Dewi Lestari Amaliah"
date: "09/01/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(brolgar)
library(naniar)
```

```{r data}
load("wages_hs2020.rda")
```


```{r wages}
wages <- as_tsibble(x = wages_hs2020,
                    key = id,
                    index = year,
                    regular = FALSE)
```


# build a linear model
```{r lm}
wages_lm <- key_slope(wages, 
                         mean_hourly_wage ~ year) %>%
  left_join(wages, by = "id")
```

```{r}

#calculate the prediction value
wages_lm <- wages_lm %>%
  mutate(pred = .intercept + (.slope_year*year))
```

```{r}
#calculate the residual
wages_res <- wages_lm %>%
  mutate(resid = mean_hourly_wage - pred)
```

```{r}

#calculate the variance of the lm model's residual 
#sort it from the highest to the lowest
wages_lm_var <- wages_res %>%
  group_by(id) %>%
  summarise(var = var(resid,
                      na.rm = TRUE)) %>%
  arrange(desc(var)) 
```


```{r}
wages_lm_var <- wages_lm_var %>%
  mutate(rank = seq(1 : nrow(wages_lm_var)))
```


### Plot the wages with high variance in their lm-residuals

```{r}
top12 <- wages_lm_var %>%
  filter(rank <=25) %>%
  left_join(wages, by = "id")

ggplot(top12, 
       aes(x = year,
           y = mean_hourly_wage)) +
  geom_miss_point() +
  geom_line() +
  facet_wrap(~id)
```

```{r}
wages %>%
  filter(id == 11974)
```




### Plot the wages with low variance in their lm-residuals
```{r}
bottom12 <- wages_lm_var %>%
  filter(rank >= 5995) %>%
  left_join(wages, by = "id")

ggplot(bottom12,
       aes(x = year,
           y = mean_hourly_wage)) +
  geom_miss_point() +
  geom_line() +
  facet_wrap(~id)
```

# take 12 random ids 
```{r}
set.seed(2021)

#take a random sample

random <- as.data.frame(sample(wages_lm_var$id, size = 30, replace = FALSE)) %>%
  rename("id" = "sample(wages_lm_var$id, size = 30, replace = FALSE)") %>%
  left_join(wages, by = "id")

ggplot(random, 
       aes(x = year,
                y = mean_hourly_wage)) +
  geom_miss_point() +
  geom_line() +
  facet_wrap(~id)

```

Learn: The observations with high and low variability were mostly the observations with the incomplete years of observations. Hence, we are going to try to only look the variance on the observations with the complete years. 


```{r}
wages_obs <- wages %>%
  features(mean_hourly_wage, n_obs) %>%
  arrange(n_obs)


complete_year_obs <- wages_obs %>%
  filter(n_obs == 27) %>%
  left_join(wages_lm_var, by = "id") %>%
  arrange(desc(var))
```

```{r}

top20_comp <- complete_year_obs %>%
  slice_head(n =20) %>%
  left_join(wages, by = "id")

ggplot(top20_comp,
       aes(x = year,
                y = mean_hourly_wage)) +
  geom_miss_point() +
  geom_line()+
  facet_wrap(~id)

```
```{r}
bottom20_comp <- complete_year_obs %>%
  slice_tail(n =20) %>%
  left_join(wages, by = "id")

ggplot(bottom20_comp,
       aes(x = year,
                y = mean_hourly_wage)) +
  geom_miss_point() +
  geom_line() +
  facet_wrap(~id)
```

```{r}
set.seed(2021)

random2 <- as.data.frame(sample(complete_year_obs$id, size = 30, replace = FALSE)) %>%
  rename("id" = "sample(complete_year_obs$id, size = 30, replace = FALSE)") %>%
  left_join(wages, by = "id")

ggplot(random2,
       aes(x = year,
                y = mean_hourly_wage)) +
  geom_miss_point() +
  geom_line() +
  facet_wrap(~id)
```

```{r}
wages %>%
  filter(year == 1984)
```

```{r}
ggplot(wages) +
  geom_boxplot(aes(x = mean_hourly_wage)) +
  facet_wrap(~year)
```

