---
title: "code_revisit"
author: "Dewi Lestari Amaliah"
date: "16/01/2021"
output: html_document
---

```{r setup, include=FALSE}
library(here)
library(readr)
library(tidyr)
library(tidyverse)
library(brolgar)
```

```{r}

# load the downloaded data 

new_data_qnames <- read_rds(here::here("data-raw/wages-high-school-demo/data-frames/wages-high-school-demo-new_data_qnames.rds"))
categories_qnames <- read_rds(here::here("data-raw/wages-high-school-demo/data-frames/wages-high-school-demo-categories_qnames.rds"))
```

# revisit the code in an observation with large variance of lm's residual (case 11974)

```{r}
get_wage_id <- function(year, id){
  new_data_qnames %>%
    select(CASEID_1979,
           paste0("HRP1_", year),
           paste0("HRP2_", year),
           paste0("HRP3_", year),
           paste0("HRP4_", year),
           paste0("HRP5_", year)
           ) %>%
    filter(CASEID_1979 == id) %>%
    rename(HRP1 = paste0("HRP1_", year),
           HRP2 = paste0("HRP2_", year),
           HRP3 = paste0("HRP3_", year),
           HRP4 = paste0("HRP4_", year),
           HRP5 = paste0("HRP5_", year))
}
```

```{r}
wage_79 <- get_wage_id(1979, 11974) 
wage_80 <- get_wage_id(1980, 11974)
wage_81 <- get_wage_id(1981, 11974)
wage_82 <- get_wage_id(1982, 11974)
wage_83 <- get_wage_id(1983, 11974)
wage_84 <- get_wage_id(1984, 11974)
wage_85 <- get_wage_id(1985, 11974)
wage_86 <- get_wage_id(1986, 11974)
wage_87 <- get_wage_id(1987, 11974)
wage_88 <- get_wage_id(1988, 11974)
wage_89 <- get_wage_id(1989, 11974)
wage_90 <- get_wage_id(1990, 11974)
wage_91 <- get_wage_id(1991, 11974)
wage_92 <- get_wage_id(1992, 11974)
wage_93 <- get_wage_id(1993, 11974)
wage_94 <- get_wage_id(1994, 11974)
wage_96 <- get_wage_id(1996, 11974)
wage_98 <- get_wage_id(1998, 11974)
wage_00 <- get_wage_id(2000, 11974)
wage_02 <- get_wage_id(2002, 11974)
wage_04 <- get_wage_id(2004, 11974)
wage_06 <- get_wage_id(2006, 11974)
wage_08 <- get_wage_id(2008, 11974)
wage_10 <- get_wage_id(2010, 11974)
wage_12 <- get_wage_id(2012, 11974)
wage_14 <- get_wage_id(2014, 11974)
wage_16 <- get_wage_id(2016, 11974)
```

```{r}

year <- c(1979, 1980, 1981, 1982, 1983, 1984, 1985, 1986, 1987, 1988, 1989, 1990, 
          1991, 1992, 1993, 1994, 1996, 1998, 2000, 2002, 2004, 2006, 2008, 2010,
          2012, 2014, 2016)
wage_11974 <- bind_rows(wage_79,
                  wage_80,
                  wage_81,
                  wage_82,
                  wage_83,
                  wage_84,
                  wage_85, 
                  wage_86,
                  wage_87,
                  wage_88,
                  wage_89,
                  wage_90,
                  wage_91,
                  wage_92,
                  wage_93,
                  wage_94,
                  wage_96,
                  wage_98,
                  wage_00,
                  wage_02,
                  wage_04,
                  wage_06,
                  wage_08,
                  wage_10,
                  wage_12,
                  wage_14,
                  wage_16) %>%
  cbind(year) %>%
  gather("raw-var", "hourly-rate", "HRP1":"HRP5", na.rm = TRUE) %>%
  arrange(year)
```

```{r}
ggplot(wage_11974) +
  geom_point(aes(x = year,
                 y = `hourly-rate`))
```

Learn: There are only 10 observations in that id and one exceptionally high wage in 1985. 


## Now, let's try for id = 11830 


```{r}

wage_79 <- get_wage_id(1979, 11830) 
wage_80 <- get_wage_id(1980, 11830)
wage_81 <- get_wage_id(1981, 11830)
wage_82 <- get_wage_id(1982, 11830)
wage_83 <- get_wage_id(1983, 11830)
wage_84 <- get_wage_id(1984, 11830)
wage_85 <- get_wage_id(1985, 11830)
wage_86 <- get_wage_id(1986, 11830)
wage_87 <- get_wage_id(1987, 11830)
wage_88 <- get_wage_id(1988, 11830)
wage_89 <- get_wage_id(1989, 11830)
wage_90 <- get_wage_id(1990, 11830)
wage_91 <- get_wage_id(1991, 11830)
wage_92 <- get_wage_id(1992, 11830)
wage_93 <- get_wage_id(1993, 11830)
wage_94 <- get_wage_id(1994, 11830)
wage_96 <- get_wage_id(1996, 11830)
wage_98 <- get_wage_id(1998, 11830)
wage_00 <- get_wage_id(2000, 11830)
wage_02 <- get_wage_id(2002, 11830)
wage_04 <- get_wage_id(2004, 11830)
wage_06 <- get_wage_id(2006, 11830)
wage_08 <- get_wage_id(2008, 11830)
wage_10 <- get_wage_id(2010, 11830)
wage_12 <- get_wage_id(2012, 11830)
wage_14 <- get_wage_id(2014, 11830)
wage_16 <- get_wage_id(2016, 11830)


```

```{r}
wage_11830 <- bind_rows(wage_79,
                  wage_80,
                  wage_81,
                  wage_82,
                  wage_83,
                  wage_84,
                  wage_85, 
                  wage_86,
                  wage_87,
                  wage_88,
                  wage_89,
                  wage_90,
                  wage_91,
                  wage_92,
                  wage_93,
                  wage_94,
                  wage_96,
                  wage_98,
                  wage_00,
                  wage_02,
                  wage_04,
                  wage_06,
                  wage_08,
                  wage_10,
                  wage_12,
                  wage_14,
                  wage_16) %>%
  cbind(year)
```

```{r}
wage_11830
```
Learn: In 1996, if we only take the mean hourly wage by the number of jobs, it means that the hourly wage for this person in 1996, would only be 2750, since the gap of wage between the first and the second job is big. Hence, should we consider to use weighted average with the number of hours work as the weight? Or looking at the history of the wage, would the 5000 wage a typo? 

However, indeed there are many extreme values of wages in that year. We can see that, in this box plot.

```{r}
load("wages_hs2020.rda")

wages_tidy <- as_tsibble(x = wages_hs2020,
                    key = id,
                    index = year,
                    regular = FALSE)
ggplot(wages_tidy) +
  geom_boxplot(aes(x = mean_hourly_wage)) +
  facet_wrap(~year)
```


# Matching the number of job in hours of work and in wage 

```{r}

get_hours_A <- function(year){
  new_data_qnames %>%
  select(CASEID_1979,
         paste0("QES-52A.01_", year),
         paste0("QES-52A.02_", year),
         paste0("QES-52A.03_", year),
         paste0("QES-52A.04_", year),
         paste0("QES-52A.05_", year)) %>%
  gather("raw-var", "hours", paste0("QES-52A.01_", year): 
           paste0("QES-52A.05_", year), 
         na.rm = TRUE) %>%
  separate("raw-var", c("mod-var", "year"), sep = -4) %>%
  separate("mod-var", c("mod-var", NA), sep = -1) %>%
  separate("mod-var", c(NA, "job-no"), sep = -1) %>%
  group_by(CASEID_1979) %>%
  summarise(total_hours = sum(hours),
         number_of_jobs = n_distinct(`job-no`)) %>%
  mutate(year = year)
}


get_hours_D <- function(year){
  new_data_qnames %>%
  select(CASEID_1979,
         paste0("QES-52D.01_", year),
         paste0("QES-52D.02_", year),
         paste0("QES-52D.03_", year),
         paste0("QES-52D.04_", year),
         paste0("QES-52D.05_", year)) %>%
  gather("raw-var", "hours", paste0("QES-52D.01_", year): 
           paste0("QES-52D.05_", year), 
         na.rm = TRUE) %>%
  separate("raw-var", c("mod-var", "year"), sep = -4) %>%
  separate("mod-var", c("mod-var", NA), sep = -1) %>%
  separate("mod-var", c(NA, "job-no"), sep = -1) %>%
  group_by(CASEID_1979) %>%
  summarise(total_hours = sum(hours),
         number_of_jobs = n_distinct(`job-no`)) %>%
  mutate(year = year)
}
```

```{r}
hours_79 <- get_hours_A(1979)
hours_80 <- get_hours_A(1980)
hours_81 <- get_hours_A(1981)
hours_82 <- get_hours_A(1982)
hours_83 <- get_hours_A(1983)
hours_84 <- get_hours_A(1984)
hours_85 <- get_hours_A(1985)
hours_86 <- get_hours_A(1986)
hours_87 <- get_hours_A(1987)
hours_88 <- get_hours_D(1988)
hours_89 <- get_hours_D(1989)
hours_90 <- get_hours_D(1990)
hours_91 <- get_hours_D(1991)
hours_92 <- get_hours_D(1992)
hours_93 <- get_hours_A(1993)
hours_94 <- get_hours_D(1994)
hours_96 <- get_hours_D(1996)
hours_98 <- get_hours_D(1998)
hours_00 <- get_hours_D(2000)
hours_02 <- get_hours_D(2002)
hours_04 <- get_hours_D(2004)
hours_06 <- get_hours_D(2006)
hours_10 <- get_hours_D(2010)
hours_12 <- get_hours_D(2012)
hours_14 <- get_hours_D(2014)
hours_16 <- get_hours_D(2016)
hours_08 <- new_data_qnames %>%
  select(CASEID_1979,
         "QES-52D.01_2008",
         "QES-52D.02_2008",
         "QES-52D.03_2008",
         "QES-52D.04_2008") %>%
  gather("raw-var", "hours", "QES-52D.01_2008":"QES-52D.04_2008", na.rm = TRUE) %>%
  separate("raw-var", c("mod-var", "year"), sep = -4) %>%
  separate("mod-var", c("mod-var", NA), sep = -1) %>%
  separate("mod-var", c(NA, "job-no"), sep = -1) %>%
  group_by(CASEID_1979) %>%
  summarise(total_hours = sum(hours),
         number_of_jobs = n_distinct(`job-no`)) %>%
  mutate(year = 2008)


hours_all_year <- bind_rows(hours_79,
                            hours_80,
                            hours_81,
                            hours_82,
                            hours_83,
                            hours_84,
                            hours_85,
                            hours_86,
                            hours_87,
                            hours_88,
                            hours_89,
                            hours_90,
                            hours_91,
                            hours_92,
                            hours_93,
                            hours_94,
                            hours_96,
                            hours_98,
                            hours_00,
                            hours_02,
                            hours_04,
                            hours_06,
                            hours_08,
                            hours_10,
                            hours_12,
                            hours_14,
                            hours_16)

```

```{r}

get_wage <- function(year){new_data_qnames %>%
  select(CASEID_1979,
           paste0("HRP1_", year),
           paste0("HRP2_", year),
           paste0("HRP3_", year),
           paste0("HRP4_", year),
           paste0("HRP5_", year)
           ) %>%
  gather("raw-var", "hourly-rate", paste0("HRP1_", year): paste0("HRP5_", year), 
         na.rm = TRUE) %>%
  separate("raw-var", c("mod-var", "year"), sep = -4) %>%
  separate("mod-var", c("mod-var", NA), sep = -1) %>%
  separate("mod-var", c(NA, "job-no"), sep = -1) %>%
  group_by(CASEID_1979) %>%
  mutate(number_of_jobs = n_distinct(`job-no`),
         mean_hourly_wage = mean(`hourly-rate`)/100) %>%    # divided by 100 to get into dollars
  filter(`job-no` == first(`job-no`)) %>%
  select(-`job-no`, -`hourly-rate`)
}

```


```{r}
wages_79 <- get_wage(1979)
wages_80 <- get_wage(1980)
wages_81 <- get_wage(1981)
wages_82 <- get_wage(1982)
wages_83 <- get_wage(1983)
wages_84 <- get_wage(1984)
wages_85 <- get_wage(1985)
wages_86 <- get_wage(1986)
wages_87 <- get_wage(1987)
wages_88 <- get_wage(1988)
wages_89 <- get_wage(1989)
wages_90 <- get_wage(1990)
wages_91 <- get_wage(1991)
wages_92 <- get_wage(1992)
wages_93 <- get_wage(1993)
wages_94 <- get_wage(1994)
wages_96 <- get_wage(1996)
wages_98 <- get_wage(1998)
wages_00 <- get_wage(2000)
wages_02 <- get_wage(2002)
wages_04 <- get_wage(2004)
wages_06 <- get_wage(2006)
wages_08 <- get_wage(2008)
wages_10 <- get_wage(2010)
wages_12 <- get_wage(2012)
wages_14 <- get_wage(2014)
wages_16 <- get_wage(2016)
```

```{r}
wages_all_year <- bind_rows(wages_79,
                  wages_80,
                  wages_81,
                  wages_82,
                  wages_83,
                  wages_84,
                  wages_85, 
                  wages_86,
                  wages_87,
                  wages_88,
                  wages_89,
                  wages_90,
                  wages_91,
                  wages_92,
                  wages_93,
                  wages_94,
                  wages_96,
                  wages_98,
                  wages_00,
                  wages_02,
                  wages_04,
                  wages_06,
                  wages_08,
                  wages_10,
                  wages_12,
                  wages_14,
                  wages_16) %>%
  rename(no_jobs = number_of_jobs) %>%
  mutate(year = as.numeric(year))
```

```{r}
# join hours and wages

hours_wages <- left_join(wages_all_year, 
                         hours_all_year, 
                         by = c("CASEID_1979", "year")) %>%
  mutate(note = ifelse(is.na(number_of_jobs) == FALSE & 
                         number_of_jobs != no_jobs, "unmatched", 
                       "matched"))

unmatched <- hours_wages %>%
  filter(note == "unmatched")
```


NOTE: when we use the weighted average of wages, there are two problems:

1. the unmatched number of jobs
2. the data of number of jobs with corresponding ID is NA




